% !TEX root = Report.tex


\section{Workflow and Project Thoughts}


	For this project, we often met together and debugged in person. More people thinking about the same thing helped things to get done more quickly. On the other hand, there were times that we focused too intensely on one idea when it would have been more useful to work in parallel on different items. For the most part however, the group worked together efficiently.
	
	Because we were such a small team, we had a few luxuries. Since we mostly worked on code in person, it was largely unnecessary to have a ticking system. Additionally, although there was some large scale division of labor (for example, Victor worked mostly on input, Louis on routing, and Moya on the system's control mechanism) there was a lot of cross-over, which was nice for when schedules became less flexible. 

	At various points in the project, we had some issues with version control. While it was useful for times when we had messed up code very badly and needed to revert, there were quite a few issues with people working on the same files. Because we were all fairly inexperienced with GIT, we figured it would be unnecessary to work on different branches. There were also multiple issues with people staging and committing more files than necessary, causing headaches with conflicts. Had members on the team been more experienced with GIT, the process would have been more streamlined. 
	
	Overall, while the team was decent with staying on deadlines, especially towards the second half of the term, there were a few weeks where progress was slow because of other committments. However, given how we had operational code for the project, we were still fairly good at staying on task. 
